\includegraphics*[width=0.4\textwidth]{clarkson.png}\\ \vspace{24pt} Wireless Security

### 2014-12-12 paper.tex
updating paper

\title{
	\includegraphics*[width=0.4\textwidth]{clarkson.png}\\
	\vspace{24pt}
	Wireless Security
}
\subtitle{
	Course\\
	Cryptography\\
	Fall 2014
}
\author{
	Virginia Weidhaas\\[12pt]
    Jihuang Hume\\[12pt]
	Alfredo Chapa\\[12pt]	
	Supervisor:\\
	Christino Tamon
}
\date{\today}

\begin{document}
\maketitle
\tableofcontents
\newpage
\section{Abstract}
This paper focuses on home WiFi security.
\newpage

\marginpar{VW}
\section{Common misbeliefs about Wifi-Security}

@@

The main difference between wired and wireless networks is the medium it transfers its data through. This difference made the burden of securing the network heavier. The broadcast nature of wireless networks makes it easy for everyone to attack the network if not secured, due to the absence of physical barriers, where the range of wireless transmission ranges from 300 ft. to half a mile \[Arbaugh2003\].

@@

In this type of attacks the attacker uses the statistics of network connectivity and activity to find information about the attacked network. Information includes: AP location, AP SSID and the type of protocol used by the analysis of size and types of packets \[Welch2003\].

@@

This type of attack is also known by many other names, such as war driving, war walking, and war flying \[Earle2005\]. This is the most common attack type where the attacker tries to get access to a network that she is not authorized to access. Mainly the reason behind such attacks is just to get Internet access for free \[Potter2003\].

@@

In this attack, the attacker gets the packets before the intended receiver does. This allows her to change the content of the message. One of the most known subset of this attack is called ARP (Address Resolution Protocol) attacks, where the attacker redirects network traffic to pass through her device \[Welch2003\].

@@

DoS (Denial of Service) attacks are the hardest type of attacks to overcome. Attackers use frequency devices to send continuous noise on a specific channel to ruin network connectivity. It is known in the wireless world as RF Jamming \[Welch2003\].

@@

\subsection{Wired Equivalent Privacy (WEP)}
Wep is the original encryption protocol developed for wireless networks. As its name implies, WEP was designed to provide the same level of security as wired networks. However, WEP has many well-known security flaws, is difficult to configure, and is easily broken.

@@

\subsection{Attacks of WEP}

Fluhrer, Mantin, and Shamir\[2002\] present several weaknesses in the key scheduling algorithm of RC4, and describe their cryptanalytic significance. They identify a large number of weak keys, in which knowledge of a small number of key bits suffices to determine many state and output bits with non-negligible probability. They use these weak keys to construct new distinguishers for RC4, and to mount related key attacks with practical complexities. They show that RC4 is completely insecure in a common mode of operation which is used in the widely deployed Wired Equivalent Privacy protocol (WEP, which is part of the 802.11 standard), in which a fixed secret key is concatenated with known IV modifiers in order to encrypt different messages. Our new passive ciphertext-only attack on this mode can recover an arbitrarily long key in a negligible amount of time which grows only linearly with its size, both for 24 and
128 bit IV modifiers.

Adam Stubblefield et al. implemented Fluhrer, Mantin, and Shamir’s attack \[FMS2002\] against WEP, the link-layer security protocol for 802.11 networks. They were able to recover the 128 bit secret key used with a passive attack. The WEP standard uses RC4 IVs improperly, and the attack exploits this design failure. They make some optimizations to make the attack more efficient, conclude that 802.11 WEP is totally insecure.

Itsik Mantin revisit a known but ignored weakness of the RC4 keystream generator, where secret state info leaks to the generated keystream, and show that this leakage, also known as Jenkins’ correlation or the RC4 glimpse, can be used to attack RC4 in several modes. Their main result is a practical key recovery attack on RC4 when an IV modifier is concatenated to the beginning of a secret root key to generate a session key. As opposed to the WEP attack from \[FMS2002\] the new attack is applicable even in the case where the first 256 bytes of the keystream are thrown and its complexity grows only linearly with the length of the key. In an exemplifying parameter setting the attack recovers a 16-byte key in 248 steps using 217 short keystreams generated from different chosen IVs. A second attacked mode is when the IV succeeds the secret root key. They mount a key recovery attack that recovers the secret root key by analyzing a single word from 222 keystreams generated from different IVs, improving the attack from \[FMS2002\] on this mode. A third result is an attack on RC4 that is applicable when the attacker can inject faults to the execution of RC4. The attacker derives the internal state and the secret key by analyzing 214 faulted keystreams generated from this key.

Tews\[2007\] demonstrated an active attack on the WEP protocol that is able to recover a 104-bit WEP key using less than 40,000 frames with a success probability of 50\%. In order to succeed in 95\% of all cases, 85,000 packets are needed. The IV of these packets can be randomly chosen. This is an improvement in the number of required frames by more than an  order of magnitude over the best known key-recovery attacks for WEP.
On an IEEE 802.11g network, the number of frames required can be obtained by re-injection in less than a minute. The required computational effort is approximately 220 RC4 key setups, which on current desktop and laptop CPUs is negligible.

\subsection{How it works}
\subsection{Flaws}
\subsection{Attacks on WEP}

\subsubsection{FMS attack}
FMS was released by Fluhrer, Mantin and Shamir in 2001 \cite{fms} and is a statistical attack on WEP.
The assumption is, that the attacker knows the IV (three bytes of the per packet key) and it uses weaknesses in RC4. Therefore he can reconstruct the key from a number of unencrypted messages.
The attack is used in tools such as AirSnort, webplab and aircrack with which it is possible to crack WEP networks.

@@

\[Arbaugh2003\] "Wireless security is different,” Computer Volume 36, Issue 8, Aug. 2003 Page(s):99 - 101\\
\[Welch2003\] "Wireless security threat taxonomy,” Information Assurance Workshop, 2003. IEEE Systems, Man and Cybernetics Society 18-20 June 2003 Page(s):76 - 83\\
\[Earle2005\] "Wireless Security Handbook,” Auerbach Publications 2005\\
\[Potter2003\] "Wireless security's future,” Security \& Privacy Magazine, IEEE Volume 1, Issue 4, July-Aug. 2003 Page(s):68 – 72\\
\[FMS2002\] Scott R. Fluhrer, Itsik Mantin and Adi Shamir: Weaknesses in the Key Scheduling Algorithm of RC4. Selected Areas in Cryptography 2001: 1-24\\
\[Tews2007\] E Tews, RP Weinmann, A Pyshkin: Breaking 104 bit WEP in less than 60 seconds - Information Security Applications, 2007 - Springer\\

@@

\bibitem{fms}
 Weaknesses in the Key Scheduling Algorithm of RC4.Shamir.

@@

\end{thebibliography}
\end{document}

>>>>

\title{
	\includegraphics*[width=0.4\textwidth]{clarkson.png}\\
	\vspace{24pt}
	Wireless Security
}
\subtitle{
	Course\\
	Cryptography\\
	Fall 2014
}
\author{
	Virginia Weidhaas\\[12pt]
    Jiaju Huang\\[12pt]
	Alfredo Chapa\\[12pt]	
	Supervisor:\\
	Christino Tamon
}
\date{\today}

\begin{document}
\maketitle
\tableofcontents
\newpage
\section{Abstract}
This paper deals with the topic WiFi Security. First some not helpful security measures are are considered, and then it goes on to common attacks on WiFis.\\
The best security measure for WiFis is a string protocol like WPA2 and a strong key\\
Therefore,the protocol WEP is being presented and discussed which flaws it has and what kind of attacks are possible on it.
In addition to this, the protocol WPA is considered and its successor WPA2.
\newpage

\section{Common misbeliefs about Wifi-Security}

@@

The main difference between wired and wireless networks is the medium it transfers its data through. This difference made the burden of securing the network heavier. The broadcast nature of wireless networks makes it easy for everyone to attack the network if not secured, due to the absence of physical barriers, where the range of wireless transmission ranges from 300 ft. to half a mile \cite{arbaugh}.

@@

In this type of attacks the attacker uses the statistics of network connectivity and activity to find information about the attacked network. Information includes: AP location, AP SSID and the type of protocol used by the analysis of size and types of packets \cite{welch}.

@@

This type of attack is also known by many other names, such as war driving, war walking, and war flying \cite{earle}. This is the most common attack type where the attacker tries to get access to a network that she is not authorized to access. Mainly the reason behind such attacks is just to get Internet access for free \cite{potter}.

@@

In this attack, the attacker gets the packets before the intended receiver does. This allows her to change the content of the message. One of the most known subset of this attack is called ARP (Address Resolution Protocol) attacks, where the attacker redirects network traffic to pass through her device \cite{welch}.

@@

DoS (Denial of Service) attacks are the hardest type of attacks to overcome. Attackers use frequency devices to send continuous noise on a specific channel to ruin network connectivity. It is known in the wireless world as RF Jamming \cite{welch}.
\\

@@

\subsection{Wired Equivalent Privacy (WEP)}
WEP was the original encryption protocol developed for wireless networks. As its name implies, WEP was designed to provide the same level of security as wired networks. However, WEP has many well-known security flaws, is difficult to configure, and is easily broken.

@@

\subsubsection{Attacks of WEP}

Fluhrer, Mantin, and Shamir\cite{fms2002} present several weaknesses in the key scheduling algorithm of RC4, and describe their cryptanalytic significance.
 They identify a large number of weak keys, in which knowledge of a small number of key bits suffices to determine many state and output bits with non-negligible probability. They use these weak keys to construct new distinguishers for RC4, and to mount related key attacks with practical complexities.\\
  They show that RC4 is completely insecure in a common mode of operation which is used in the widely deployed Wired Equivalent Privacy protocol (WEP, which is part of the 802.11 standard), in which a fixed secret key is concatenated with known IV modifiers in order to encrypt different messages.

That new passive ciphertext-only attack on this mode can recover an arbitrarily long key in a negligible amount of time which grows only linearly with its size, both for 24 and
128 bit IV modifiers.\\
Adam Stubblefield et al. implemented Fluhrer, Mantin, and Shamir’s attack \cite{fms2002} against WEP, the link-layer security protocol for 802.11 networks. They were able to recover the 128 bit secret key used with a passive attack.

The WEP standard uses RC4 IVs improperly, and the attack exploits this design failure. They make some optimizations to make the attack more efficient, conclude that 802.11 WEP is totally insecure.

Itsik Mantin revisit a known but ignored weakness of the RC4 keystream generator, where secret state info leaks to the generated keystream, and show that this leakage, also known as Jenkins’ correlation or the RC4 glimpse, can be used to attack RC4 in several modes.

Their main result is a practical key recovery attack on RC4 when an IV modifier is concatenated to the beginning of a secret root key to generate a session key.

As opposed to the WEP attack from \cite{fms2002} the new attack is applicable even in the case where the first 256 bytes of the keystream are thrown and its complexity grows only linearly with the length of the key.

In an exemplifying parameter setting the attack recovers a 16-byte key in 248 steps using 217 short keystreams generated from different chosen IVs. A second attacked mode is when the IV succeeds the secret root key. They mount a key recovery attack that recovers the secret root key by analyzing a single word from 222 keystreams generated from different IVs, improving the attack from \cite{fms2002} on this mode.

A third result is an attack on RC4 that is applicable when the attacker can inject faults to the execution of RC4. The attacker derives the internal state and the secret key by analyzing 214 faulted keystreams generated from this key.\\

Tews \cite{tews} demonstrated an active attack on the WEP protocol that is able to recover a 104-bit WEP key using less than 40,000 frames with a success probability of 50\%. In order to succeed in 95\% of all cases, 85,000 packets are needed. The IV of these packets can be randomly chosen. This is an improvement in the number of required frames by more than an  order of magnitude over the best known key-recovery attacks for WEP.
On an IEEE 802.11g network, the number of frames required can be obtained by re-injection in less than a minute. The required computational effort is approximately 220 RC4 key setups, which on current desktop and laptop CPUs is negligible.

\subsubsection{FMS attack}
FMS was released by Fluhrer, Mantin and Shamir in 2001 \cite{fms2002} and is a statistical attack on WEP.
The assumption is, that the attacker knows the IV (three bytes of the per packet key) and it uses weaknesses in RC4. Therefore he can reconstruct the key from a number of unencrypted messages.
The attack is used in tools such as AirSnort, webplab and aircrack with which it is possible to crack WEP networks.

@@

@@

@@

\bibitem{arbaugh}
"Wireless security is different,” Computer Volume 36, Issue 8, Aug. 2003 Page(s):99 - 101\\

\bibitem{welch}
"Wireless security threat taxonomy,” Information Assurance Workshop, 2003. IEEE Systems, Man and Cybernetics Society 18-20 June 2003 Page(s):76 - 83\\

\bibitem{earle}
"Wireless Security Handbook,” Auerbach Publications 2005\\

\bibitem{potter}
"Wireless security's future,” Security \& Privacy Magazine, IEEE Volume 1, Issue 4, July-Aug. 2003 Page(s):68 – 72\\

\bibitem{fms2002}
 Scott R. Fluhrer, Itsik Mantin and Adi Shamir: Weaknesses in the Key Scheduling Algorithm of RC4. Selected Areas in Cryptography 2001: 1-24\\

\bibitem{tews} E Tews, RP Weinmann, A Pyshkin: Breaking 104 bit WEP in less than 60 seconds - Information Security Applications, 2007 - Springer\\
\end{thebibliography}
\end{document}
